---
title: "Creating a Simple Interactive Map with R and Leaflet"
author: "Neil Kutty"
date: "September 1, 2016"
output: html_document
---

deployed Shiny app: [Crime DC - Last 30 Days](https://kuttyman.shinyapps.io/crimeDC)

data: DC Crime - Last 30 Days [Open Data DC](http://opendata.dc.gov/datasets)

***

### Load needed libraries.

  + `dplyr` for distinct(), select(), and mutate() functions.
  + `tidyr` for the separate() function.
  + `jsonlite` for the fromJSON() function needed to retrieve our dataset via the GeoJSON api.
  + `lubridate` for ymd_hms() function to clean datetime column.
  + `leaflet` to create our map.


```{r setup, include=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(jsonlite)
library(lubridate)
library(leaflet)
```



### Clean and transform data from GeoJSON format into a dataframe

**Fig 1.**

Our data comes in a `GeoJSON` format.  For a good primer on dealing with this and `JSON` data sources in general, see: 

 + [GeoJSON Format](https://tools.ietf.org/html/rfc7946)
 + [Using jsonlite library in R](https://cran.r-project.org/web/packages/jsonlite/vignettes/json-aaquickstart.html)

The data source link can be viewed/downloaded [here](http://opendata.dc.gov/datasets/dc3289eab3d2400ea49c154863312434_8.geojson). Note that it is in JSON document/collection format.  The `jsonlite::fromJSON()` function retrieves the from the collection format it exists in and stores it in list format in the variable defined `dccrimejsonlite`.

We access the list elements which are nested within the `features` level, and then combine these two large lists:

- `properties` list which contain the main table elements for the dataset. 
- `geometry` list which contains the **latitude** and **longitude** columns which we will need to map the data. 

Using the `cbind` function, we combine these two lists into a dataframe defined as `dc_crime_json`.  Our next step is to clean the data and return a usable dataframe.  

```{r cleandata, eval=T, message=F, warning=F}
# Fig. 1 

########---------------------------------------------------------------------#>>>
  ## Retrieve the data in JSON format from opendata.dc.gov using fromJson()
  dccrimejsonlite <- fromJSON('http://opendata.dc.gov/datasets/dc3289eab3d2400ea49c154863312434_8.geojson')
  ## use cbind() combine the list elements and create a dataframe
  dc_crime_json <- cbind(dccrimejsonlite$features$properties,dccrimejsonlite$features$geometry)

  ## Seperate and clean lat/long columns but keep original datetime column
  ## --also separate REPORTDATETIME column
  dc_crime_clean <- dc_crime_json %>% 
    separate(coordinates, into = c("X", "Y"), sep = ",")%>%
    separate(REPORT_DAT, into = c("Date","Time"), sep="T", remove = FALSE)%>%
    mutate(Weekday = weekdays(as.Date(REPORT_DAT)),
           DATETIME = ymd_hms(REPORT_DAT, tz='America/New_York'),
           Date = as.Date(Date),
           X = as.numeric(gsub("c\\(","",X)),
           Y = as.numeric(gsub("\\)","",Y)))
  
  print(head(dc_crime_clean,n=4))

```


### Render the leaflet map

First, we define a `points` element with the latitude and longitude vectors corresponding to the dataframe. This is used to pass the Latitude and Longitude columns to the leaflet map: see below where `addMarkers(data = points ...`.

Rendering a map with leaflet is as easy as calling the `leaflet()` function.  Using chain operation, we simply add elements to our map after calling it.  


```{r map, eval=F}
#This code does not run, please see Shiny App for Map Render
points <- cbind(dc_crime_clean$X,dc_crime_clean$Y)
leaflet() %>%
  addProviderTiles("OpenStreetMap.Mapnik",
                   options = providerTileOptions(noWrap = TRUE)
  ) %>%
  addMarkers(data = points,
             popup = paste0("<strong>Report Date: </strong>",
                            dc_crime_clean$DateClean,
                            "<br><strong>Offense: </strong>", 
                            dc_crime_clean$OFFENSE, 
                            "<br><strong>method: </strong>", 
                            dc_crime_clean$METHOD,
                            "<br><strong>shift: </strong>",
                            dc_crime_clean$SHIFT,
                            "<br><strong>blocksite address: </strong><br>",
                            dc_crime_clean$BLOCKSITEADDRESS
             ),
             clusterOptions = markerClusterOptions()
  ) 
  
```


