---
title: "Creating a Simple Interactive Map with R and Leaflet"
author: "Neil Kutty"
date: "September 1, 2016"
output: html_document
---

deployed Shiny app: [Crime DC - Last 30 Days](https://kuttyman.shinyapps.io/crimeDC)

data: DC Crime - Last 30 Days [Open Data DC](http://opendata.dc.gov/datasets)

***

### Load needed libraries.

  + `dplyr` for distinct(), select(), and mutate() functions.
  + `tidyr` for the separate() function.
  + `jsonlite` for the fromJSON() function needed to retrieve our dataset via the GeoJSON api.
  + `lubridate` for ymd_hms() function to clean datetime column.
  + `leaflet` to create our map.

```{r setup, include=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(jsonlite)
library(lubridate)
library(leaflet)
```



### Clean and transform data from GeoJSON format into a dataframe

**Fig 1.**

Our data comes in a `GeoJSON` format.  For a good primer on dealing with this and `JSON` data sources in general, see: 

 + [GeoJSON Format](https://tools.ietf.org/html/rfc7946)
 + [Using jsonlite library in R](https://cran.r-project.org/web/packages/jsonlite/vignettes/json-aaquickstart.html)

The data source link can be viewed/downloaded [here](http://opendata.dc.gov/datasets/dc3289eab3d2400ea49c154863312434_8.geojson). Note that it is in JSON document/collection format.  The `jsonlite::fromJSON()` function retrieves the from the collection format it exists in and stores it in list format in the variable defined `dccrimejsonlite`.

We access the list elements which are nested within the `features` level, and then combine these two large lists:

- `properties` list which contain the main table elements for the dataset. 
- `geometry` list which contains the **latitude** and **longitude** columns which we will need to map the data. 

Using the `cbind` function, we combine these two lists into a dataframe defined as `dc_crime_json`.  Our next step is to clean the data and return a usable dataframe.  

```{r cleandata, eval=TRUE}
# Fig. 1 

########---------------------------------------------------------------------#>>>
  ## Retrieve the data in JSON format from opendata.dc.gov using fromJson()
  dccrimejsonlite <- fromJSON('http://opendata.dc.gov/datasets/dc3289eab3d2400ea49c154863312434_8.geojson')
  ## use cbind() combine the list elements and create a dataframe
  dc_crime_json <- cbind(dccrimejsonlite$features$properties,dccrimejsonlite$features$geometry)

  ## Seperate and clean lat/long columns but keep original datetime column
  ## --also separate REPORTDATETIME column
  dc_crime_clean <- dc_crime_json %>% 
    separate(coordinates, into = c("X", "Y"), sep = ",")%>%
    separate(REPORTDATETIME, into = c("Date","Time"), sep="T", remove = FALSE)%>%
    mutate(Weekday = weekdays(as.Date(REPORTDATETIME)),
           DATETIME = ymd_hms(REPORTDATETIME, tz='America/New_York'),
           Date = as.Date(Date),
           X = as.numeric(gsub("c\\(","",X)),
           Y = as.numeric(gsub("\\)","",Y)))

```


### Render the leaflet map

First, we define a `points` element with the latitude and longitude vectors corresponding to the dataframe. This is used to pass the Latitude and Longitude columns to the leaflet map: see below where `addMarkers(data = points ...`.

Rendering a map with leaflet is as easy as calling the `leaflet()` function.  Using chain operation, we simply add elements to our map after calling it.  


```{r map, eval=TRUE}
points <- cbind(dc_crime_clean$X,dc_crime_clean$Y)
leaflet() %>%
  addProviderTiles("OpenStreetMap.Mapnik",
                   options = providerTileOptions(noWrap = TRUE)
  ) %>%
  addMarkers(data = points,
             popup = paste0("<strong>Report Date: </strong>",
                            dc_crime_clean$DateClean,
                            "<br><strong>Offense: </strong>", 
                            dc_crime_clean$OFFENSE, 
                            "<br><strong>method: </strong>", 
                            dc_crime_clean$METHOD,
                            "<br><strong>shift: </strong>",
                            dc_crime_clean$SHIFT,
                            "<br><strong>blocksite address: </strong><br>",
                            dc_crime_clean$BLOCKSITEADDRESS
             ),
             clusterOptions = markerClusterOptions()
  ) 
  
```


#### Below are the full ui and server files for the Shiny app

###ui.R 
```{r ui, eval=FALSE}
library(shiny)
library(shinythemes)
library(leaflet)

fluidPage(#theme = shinytheme("cerulean"),
          tags$head(
            # Include our custom CSS
           includeCSS("style.css")
           
          ),
          navbarPage("Crime DC",id='nav',
            tabPanel("Interactive Map",
                     h5('These charts update to reflect data points in the current viewable area of the map. 
                        The Data Explorer tab shows the corresponding rows of the viewable area.'),
                 
                     
                       
                     fluidRow(column(7,leafletOutput("mymap", width = '100%', height = '400px')),
                              column(5,plotOutput("plotDayTime"))),
                     
                     
                     fluidRow(
                              column(4,plotOutput("plotOffense")),
                              column(4,plotOutput("plotTimeline")),
                              column(4,plotOutput("plotSeries"))
                              ),
                     
                    
                      
                     br(),
                     absolutePanel(id = "controls",class = "panel panel-default", fixed = TRUE, draggable = TRUE,
                                   top = 225, left = 50, right = "auto", bottom = "auto",
                                   width = 150, height = "auto", style = "opacity: .65",
                                           
                                   h3("Crime DC - Last 30 Days"), br(), 
                                   actionButton("resetMap", "Reset Map", style="color: #fff; background-color: #337ab7; border-color: #2e6da4"),
                                   br(),
                                   paste("Click on cluster or scroll to zoom-in, Click an individual marker for additional detail popup.")
                    )
          ),
          tabPanel("Data Explorer",
                   dataTableOutput("table1")
                   
          )
        ),
        a("data source: http://opendata.dc.gov/datasets",href="http://opendata.dc.gov/datasets"),
        paste(' | '),
        a("direct data source api link:",href='http://opendata.dc.gov/datasets/dc3289eab3d2400ea49c154863312434_8.geojson'),
        paste(' | '),
        a("author: neil kutty", href="http:/twitter.com/neilkutty"),
        paste(' | '),
        a("github",href="https://github.com/sampsonsimpson/DC_Crime_Data"),
        paste(' | '),
        a("neighborhood boundaries KML file",href="https://www.google.com/maps/d/viewer?mid=1z_3yTY-G8hZZ3z5qh3tM9dBh5ps&hl=en_US")
        
)


```

***

###server.R
```{r server.R file, eval=FALSE}

library(ggplot2)
library(leaflet)
library(dplyr)
library(tidyr)
library(jsonlite)
library(curl)
library(lubridate)
library(rgdal)



########---------------------------------------------------------------------#>>>
  ## Retrieve the data in JSON format from opendata.dc.gov using fromJson()
  dccrimejsonlite <- fromJSON('http://opendata.dc.gov/datasets/dc3289eab3d2400ea49c154863312434_8.geojson')
  ## use cbind() combine the list elements and create a dataframe
  dc_crime_json <- cbind(dccrimejsonlite$features$properties,dccrimejsonlite$features$geometry)
  
  ## Get distinct Offenses for shiny input
  offenses <- distinct(select(dc_crime_json,OFFENSE))
  row.names(offenses) <- offenses$OFFENSE
  
  ## Seperate and clean lat/long columns but keep original datetime column
  ## --also separate REPORTDATETIME column
  dc_crime_clean <- dc_crime_json %>% 
    separate(coordinates, into = c("X", "Y"), sep = ",")%>%
    separate(REPORTDATETIME, into = c("Date","Time"), sep="T", remove = FALSE)%>%
    mutate(Weekday = weekdays(as.Date(REPORTDATETIME)),
           Date = as.Date(Date),
           X = as.numeric(gsub("c\\(","",X)),
           Y = as.numeric(gsub("\\)","",Y)))
  dc_crime_clean$DATETIME = as.POSIXct(strptime(dc_crime_clean$REPORTDATETIME, tz = "UTC", "%Y-%m-%dT%H:%M:%OSZ"))  
  
  dc_crime_lite <- dc_crime_clean %>%
    select(OFFENSE,SHIFT,DATETIME,Date,X,Y,Weekday,BLOCKSITEADDRESS,METHOD)
           
 
  dchoods <- readOGR("dchoods.kml", "DC neighborhood boundaries")
  
#Shiny server
function(input, output, session) {
    
    filterData <- reactive({
      if (is.null(input$mymap_bounds))
      return(dc_crime_lite)
      bounds <- input$mymap_bounds
      latRng <- range(bounds$north, bounds$south)
      lngRng <- range(bounds$east, bounds$west)
      filter(dc_crime_lite, Y >= latRng[1] & Y <= latRng[2] & X >= lngRng[1] & X <= lngRng[2])
  })

  output$plotOffense <-  
    if(is.null(filterData)){
    return()
    }else{
      renderPlot({
      off <- as.data.frame(table(filterData()$OFFENSE))
      off$Freq <- as.numeric(off$Freq)
      off$Var1 <- factor(off$Var1)
      colnames(off) <- c("OFFENSE","COUNT")
      ggplot(off, aes(x=OFFENSE,y=COUNT)) +
        geom_bar(stat="identity",alpha = 0.45, fill='red') +
        ggtitle("Number of Crimes by Offense") +
        geom_text(aes(label = off$COUNT), size = 3.5, hjust = .58, color = "black")+
        coord_flip()+
        scale_x_discrete(label = function(x) lapply(strwrap(x, width = 10, simplify = FALSE), paste, collapse="\n"))+
        theme(axis.title=element_text(size=10),
              axis.text.y = element_text(size=10, hjust = 1),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              panel.border = element_blank(),
              panel.background = element_blank()
              
              )
      
      })}

output$plotSeries <-
    if(is.null(filterData)){
      return()
    }else{
      renderPlot({
        ts <- filterData() %>%
          select(Date) %>%
          group_by(Date) %>%
          summarize(count = n())
        
        ggplot(ts[-c(nrow(ts),1),], aes(x=Date, y=count, alpha = 0.8))+
          geom_line()+
          geom_text(aes(label = ts[-c(nrow(ts),1),]$count), size = 3.5, hjust = .58, color = "black")+
          ggtitle("Number of Crimes by Day")+
          guides(alpha=FALSE)+
          theme(axis.title=element_text(size=10),
                axis.text.x = element_text(size = 10, angle = 45, hjust = 1))+
          theme_bw()
      })
    }
  
    
output$plotTimeline <-
  if(is.null(filterData)){
    return()
  }else{
    renderPlot({
      scat <- filterData() %>%
        select(OFFENSE, Date) %>%
        group_by(OFFENSE, Date) %>%
        summarize(count = n())
      
      ggplot(scat, aes(x=Date, y=count, color=OFFENSE))+
        geom_point()+
        ggtitle("Number of Crimes by Day by Offense")+
        scale_fill_brewer("Set2")+
        theme(axis.title=element_text(size=10),
              axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
              panel.background = element_rect(fill = "white"),
              strip.background = element_rect(fill = "white"),
              legend.position = c(.35,.75),
              legend.background = element_rect(fill=alpha('white', 0.2)))
    })
  }

  
  output$plotDayTime <-
    if(is.null(filterData)){
      return()
    }else{
    renderPlot({
      dt <- filterData() %>%
        select(Weekday, SHIFT) %>%
        group_by(Weekday, SHIFT) %>%
        summarize(count = n())
      dt$Weekday <- factor(dt$Weekday, levels= c("Sunday", "Monday","Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
      dt$SHIFT <- factor(dt$SHIFT, levels= c("DAY","EVENING","MIDNIGHT"))
      dt[order(dt$Weekday,dt$SHIFT),]
      ggplot(dt,aes(x=SHIFT,y=count,fill=SHIFT)) +
        geom_bar(stat="identity", alpha = 0.75) +
        scale_fill_brewer(palette = 'Set2')+
        scale_y_continuous()+
        ggtitle("Number of Crimes by Day of Week and Time of Day (SHIFT)")+
        facet_grid(.~Weekday)+
        theme(axis.title=element_text(size=10),
              axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
              panel.background = element_rect(fill = "white"),
              strip.background = element_rect(fill = "white"),
              legend.position = c(.085,.9),
              legend.background = element_rect(fill=alpha('white', 0.2)))
      })}
      
   output$table1 <- 
     renderDataTable(options=list(pageLength=25),{
       filterData()%>%
         select(Weekday, SHIFT, DATETIME, BLOCKSITEADDRESS, OFFENSE, METHOD)
     })
  
   points <- eventReactive(input$resetMap,{
     
     cbind(dc_crime_clean$X,dc_crime_clean$Y)
     
   }, ignoreNULL = FALSE)

   output$mymap <- renderLeaflet({
    
    leaflet() %>%
      addProviderTiles("OpenStreetMap.Mapnik",
                       options = providerTileOptions(noWrap = TRUE)
      ) %>%
      addMarkers(data = points(),
                 popup = paste0("<strong>Report Date: </strong>",
                                dc_crime_clean$DATETIME,
                                "<br><strong>Offense: </strong>", 
                                dc_crime_clean$OFFENSE, 
                                "<br><strong>method: </strong>", 
                                dc_crime_clean$METHOD,
                                "<br><strong>shift: </strong>",
                                dc_crime_clean$SHIFT,
                                "<br><strong>blocksite address: </strong><br>",
                                dc_crime_clean$BLOCKSITEADDRESS
                 ),
                 clusterOptions = markerClusterOptions()
      ) %>%
       addPolygons(data = dchoods, 
                   fillOpacity = 0.2, 
                   color = 'blue',
                   fillColor = 'white',
                   weight = 2.0
                  )
    
  })
}
```

